---
title: "Overview gameday"
author: "Santina Lin"
date: "Nov 17, 2014"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of Gameday}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

This package extracts information from the NHL server and provide the information the user needs using either `gday()` or `score()`. The former returns the broadcaster name if the team is playing, and the latter returns a data frame containing the information about the game (if given) or all the games on the current day. 

More functions will be added to this package (once I have more time), which will allow more functionalities and processing of the data queried from the NHL server. 

This is also my first R package. Though minimum but fully functional and a great learning opportunity. 

## Using the functions 

Both `gday()` and `score()` have default parameters. `gday()` assumes its parameter `team` is "canucks" if no input is given. `score()` takes in two parameters, `team` and `date`, in such order. 

The default value for the date for both functions is the current date (`Sys.Date()`) and the team is set to default. 

In `score()` When `team` is not given as an input, the function returns all the games playing on that date. If `team` is given, it will check if it is in the list of the games on that date and return the game information. If the input team is not playing, it will simply return a data.frame of all the games on the current day.

Some examples: 
```{r}
library(gameplay)
gday(date = "2014-11-18", team = "KINGS") # returns the broadcaster name for Oilers 
gday(team = "Dubai") # return "not playing" since Dubai isn't in the NHL
gday(date = "2014-11-19") # check for Canucks, which is playing on Nov 19, 2014
```

## Error handling 

The date format has to be in the specific format of `YYYY-MM-DD`, and need to be in numbers and in quote.  This check is performed at the start of both functions `gday()` and `score()` to ensure proper handling of the inputs. 
If the date is not an acceptable format, it will throw one of the two different errors.

Examples (try running the commented-out code)
```{r}
# detects improper length 
#gday(date = "2014-111-18") 
# detects not all date components are numbers
#score(date = "2014-No-18") 
```


And if date such as "1111-12-31", `gday()` and `score()` will handle the situation when NHL server returns 404 error. 
Examples: 
```{r}
gday(date = "2022-11-18") 
score(date = "1700-11-18")
```


## Challenges I found

Though this first version of the package seems short, it took me forever to figure out some of the really important thing in building R package. I'm listing them out here as notes to (future) self and perhaps others. 

### devtools::check() is independent of your environment 
That said, it could throw you  errors that you otherwise would not see because of the libraries exist in your current R studio environment has enabled you to run your functions. So for example, if you don't specify the libraries such as `RCurl` in DESCRIPTION, it would not know what `getURL()` function is. 

So you need to add `RCurl` under `Imports` (there's an s) in your DESCRIPTION. Moreover, make sure you specify the library when calling any non-base function. I.e., always say something like `RCurl::getURL()` instead of just `getURL()` in your exported functions. 

### Libraries in DESCRIPTION 
A common practice is to have all the libraries that are required to run the functions in the `R` directory __and knitting the vignettes markdown__ in the package directory, under `Imports` in DESCRIPTION. 

Under `Depends` it is usually just the version of the R. 

[Correct me](santina424@gmail.com) if I'm wrong. 

### Helper functions 
The R scripts that act as helper functions to the exported functions don't need to have documentation like the exported functions do. If you do have documentations that prompt `document()` to generate .Rd files for them, don't have example at the end since `devtools::test()` would complain that they're not found (since they're not exported). Only exported functions need documentations and examples. 

If you really want to "document" the helper functions, just use `#` instead of `#'` to write comments. 


## Reflection

This was overall a good learning experience. I felt like I didn't write as much code given the time I spent on it... Most of the time were spent on understanding the workflow of building R packages and the magic behinds the functions in devtools. 

Given now I know how to do this, I think it'd be helpful to write my own functions and include them in a package for repetitive use, such as in a research setting or whatever.  For example, if the next 5 assignments are about graphing gapminde, I'd write a package that is designed specifically for spitting out graphs. 

It's also great that we have sorted seen how getting informatino from the web works. 
